---
title: "mvabund model selection"
author: "JBW"
date: "12/5/2021"
output: pdf_document
---

***GLM: Include depth, but not site or their interaction
***For PCA, get statistic (using envfit, adonis)
***Univariate GLMs for each group, w/ model selection

First, load the libraries
```{r}
library(zoo) # as.Date
library(mvabund) #manyglm, etc
library(vegan) #rda
library(ggplot2)
library(ggfortify)
library(ggpubr) #ggarrange
library(gridExtra) ##grid.arrange
library(grid) #textGrob
library(reshape2)
```

Then, load the data
```{r}
# Load data
data = read.csv("data/combined_counts.csv")
data$Date = as.Date(data$Date,"%m/%d/%y")
data$Depth_ctgry = factor(data$Depth_ctgry)

# Pull out only the columns of plankton counts
plankton_counts = data[,8:18]; plankton_counts$copepods = data$copepods

# Make smaller data set omitting 3 rarest groups (fish, crabs, shrimp)
plankton_common = cbind(plankton_counts[,2:6],plankton_counts[,9:12])

# Create mvabund object for common plankton groups
plankton_mv = mvabund(plankton_common)
```

To avoid redundant predictor terms, run quick PCA on physical data
```{r}
data_waterqual = as.data.frame(cbind(data$DO,data$temp_C,data$turbidity,
                                data$salinity,data$chlorophyll,data$DOM,data$BGA))
colnames(data_waterqual) = c("DO","temp","turbidity","sal","chlor","DOM","BGA")
water.pca = rda(na.omit(data_waterqual), scale=T) #different usits, have to scale
#summary(water.pca)
screeplot(water.pca,type="l")
biplot(water.pca,type="text")
```
Turbidity, chlorophyll, and blue-green algae all (understandably) co-vary. So, we will only include 1 of the three in our model. We will stick with chlorophyll, since it is commonly measured in other studies.

Now, model comparison is best if run on the same dataset (AKA NA values have to be dealt with). So, we will run linear interpolation to fill in the 16 instances of NA's in chlorophyll and DOM.
```{r}
#create new object and use linear interpolation to fill NAs
data_nona = data
data_nona$chlorophyll = na.approx(data_nona$chlorophyll,method="linear",na.rm=F)
data_nona$DOM = na.approx(data_nona$DOM,method="linear",na.rm=F)
```

Now for model selection... 
First, run the full model (omitting pH due to QC concerns), allowing for all possible 2-way interactions between factors, and all possible 2-way interactions between continuous variables
```{r}
glm1 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$salinity + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity + 
                 data_nona$temp_C:data_nona$chlorophyll + 
                 data_nona$temp_C:data_nona$DOM + 
                 data_nona$salinity:data_nona$chlorophyll + 
                 data_nona$salinity:data_nona$DOM + 
                 data_nona$chlorophyll:data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI") #default
#for some reason, the above model would not run until I stopped allowing interactions between the factor variables (site, depth category) and the numeric ones...

plot(glm1) #check residuals
glm1 #get summary info
```

Then use drop1 function to test models with fewer terms
```{r}
drop_tests = drop1(glm1,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_tests
```
This suggests dropping the interaction between chlorophyll and DOM...

So run the model again, dropping the term that yielded the lowest AIC
```{r}
glm2 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$salinity + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity + 
                 data_nona$temp_C:data_nona$chlorophyll + 
                 data_nona$temp_C:data_nona$DOM + 
                 data_nona$salinity:data_nona$chlorophyll + 
                 data_nona$salinity:data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI") #default

plot(glm2) #check residuals
drop_test_2 = drop1(glm2,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_2
```
Since we still have a lot of terms, we will go ahead and toss out the interaction term with the next lowest AIC: the interaction between DO and salinity.

```{r}
glm3 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity + 
                 data_nona$temp_C:data_nona$chlorophyll + 
                 data_nona$temp_C:data_nona$DOM + 
                 data_nona$salinity:data_nona$chlorophyll + 
                 data_nona$salinity:data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI") #default

plot(glm3) #check residuals
drop_test_3 = drop1(glm3,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_3
```
2022 EDIT: NOW ALL REMAINING TERMS RESULT IN HIGHER AICs! THIS SUGGESTS THAT GLM3 IS THE BEST MODEL FOR THE COMMUNITY

```{r}
model_2022 = anova.manyglm(glm3) #took 23 min
model_2022
```

Again. since we still have a lot of terms, we will go ahead and toss out the interaction term with the next lowest AIC: the interaction between salinity and chlorophyll. >> 2022 EDIT! THE AICs WERE WORSE, BUT THE FINAL COMMUNITY MODEL STILL HAS NON-SIGNIFICANT TERMS...... SO LET'S TOSS TEMPxCHLOR

```{r}
glm4 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity +  
                 data_nona$temp_C:data_nona$DOM + 
                 data_nona$salinity:data_nona$chlorophyll + 
                 data_nona$salinity:data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI") #default

plot(glm4) #check residuals
drop_test_4 = drop1(glm4,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_4
```
At this point, no single (dropped) term can make the model significantly better. However, we will keep dropping the lowest AIC value just to see if we get to another point when the model improves again. >> 2022 EDIT: SAME REASONING, BUT *NOW* DROPPING SALxCHLOR, WHICH GETS UP BACK TO THE AIC FROM GLM3

```{r}
glm5 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity +  
                 data_nona$temp_C:data_nona$DOM +  
                 data_nona$salinity:data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI", #default
               show.coef = F) 

plot(glm5) #check residuals
drop_test_5 = drop1(glm5,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_5
```
...It only got worse... So we will stick with glm4, which kept the term dropped in glm5. >> 2022 EDIT, GLM5 LOOKS GOOD

So next, we get the univariate tests...
########
```{r}
model1 = anova.manyglm(glm4,p.uni = "adjusted") #took 26 min
model1 #get p-values
#saveRDS(model1,"results/GLM_model_12.5.21.rds")
```

In the full output, we see that the interaction between temperature and chlorophyll is not significant in either the multivariate test or in any univariate tests. Therefore, we will try the same test for glm5 (when we dropped that term and the AIC stayed the same).

```{r}
model2 = anova.manyglm(glm5) #took 19 min
#model2 = readRDS("results/GLM_model2_12.5.21.rds")
model2 #get p-values
#saveRDS(model2,"results/GLM_model2_12.5.21.rds")
```

Just out of curiousity, what happens when we drop all interaction terms...
```{r}
glm6 = manyglm(plankton_mv ~ data_nona$Site + 
                 data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI") #default

plot(glm6) #check residuals
drop_test_6 = drop1(glm6,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_6
```

2022: Let's keep dropping terms from glm5 to see what happens! Specifically, let's drop temp*DOM
```{r}
glm7 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity +  
                 data_nona$salinity:data_nona$DOM,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI", #default
               show.coef = F) 

plot(glm7) #check residuals
drop_test_7 = drop1(glm7,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_7

#summary.manyglm(glm5,show.est = T)
```
2022: The AICs are all worse, but let's see if the model still has non-significant terms...
```{r}
model3 = anova.manyglm(glm7) #took 17.5 min
model3 #get p-values
```
2022: Again, we have terms that are borderline significant (sal*DOM) so let's drop them!

```{r}
glm8 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$chlorophyll + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI", #default
               show.coef = F) 

plot(glm8) #check residuals
drop_test_8 = drop1(glm8,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_8
```
2022: The AICs are still getting worse, but let's go ahead and see whether the remaining terms are all significant.

```{r}
model5 = anova.manyglm(glm8) #took 15 min
#model2 = readRDS("results/GLM_model2_12.5.21.rds")
model5 #get p-values
#saveRDS(model2,"results/GLM_model2_12.5.21.rds")
```
2022: Out of curiousity, what happens when we ditch that DO*chlor term.....

```{r}
glm9 = manyglm(plankton_mv ~ data_nona$Depth_ctgry + 
                 data_nona$DO + 
                 data_nona$temp_C + 
                 data_nona$salinity +
                 data_nona$chlorophyll +
                 data_nona$DOM + 
                 data_nona$DO:data_nona$temp_C + 
                 data_nona$DO:data_nona$DOM + 
                 data_nona$temp_C:data_nona$salinity,
               family="negative_binomial", #better for counts
               nBoot = 1000, #default
               k=1, #default
               theta.method = "PHI", #default
               show.coef = F) 

plot(glm9) #check residuals
drop_test_9 = drop1(glm9,test="Chisq",trace=T) #likelihood ratio test recommended for GLMs
drop_test_9
```
2022: The AICs are still getting worse, but let's go ahead and see what happens...

```{r}
model6 = anova.manyglm(glm9) #took 13 min
#model2 = readRDS("results/GLM_model_01.05.22.rds")
model6 #get p-values
#saveRDS(model6,"results/GLM_model_01.05.22.rds")
```
```{r}

```

